---
title: "Introduction to R"
author: "BotnieData"
date: "2024-11-28"
output: html_document
---

### Topics
- Comment
- Basic Arithmetic
- Variable


### **Comment**
`#` symbol is used to start a comment. Any text following # on the same line will be ignored by the R interpreter. 
This is a convenient way to add explanations or notes to your code without affecting its execution.

Formatting and organizing code within the script and document by using multiple levels of headings (`#`, `##`, `###`, `####`), you can create a hierarchical structure that visually outlines different sections of your code. This can make it easier to navigate and understand, especially for larger or more complex projects.

#### *Examples:*
```{r}
# This is a comment, Calculate the sum of 2 and 2
2 + 2

# This is a comment, Calculate the difference of 27 subtracted by 10
27 - 10
```

### **Basic Arithmetic**
R is a powerful tool for performing various mathematical calculations. It supports a wide range of arithmetic operations, including:

#### **Arithmetic Operators**
- Multiplication: `*`
- Division: `/`
- Addition: `+`
- Subtraction: `-`
- Exponentiation: `^`
- Modulo: `%%`

#### *Examples:*
```{r}

# Multiplication
# This is a comment, Calculate the product of 9 multiplied by 8
9 * 8

# Division
# This is a comment, Calculate the quotient of 6 divided by 2
6 / 2

# Addition
# This is a comment, Calculate the sum of 1 and 1
1 + 1

# Subtraction
# This is a comment, Calculate the difference of 27 subtracted by 10
27 - 10

# Exponential
# # This is a comment, Calculate 'yth power of x', so called 
# "The 5th power of 2 is 32"
2^5 # (2 * 2 * 2 * 2 * 2)

# Modulo
# # This is a comment, Calculate the remainder after division
# "28 modulo 6 is 4"
28%%6 # 28/6 = 24 remainder and remainder 4 is the modulo
```

