---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: false
    theme: cerulean
---

## **Topics**
- Comment
- Basic Arithmetic
- Variable
- Data types
- `class()` function
- `str()` function

---

### **Comment**
`#` symbol is used to start a comment. Any text following # on the same line will be ignored by the R interpreter. 
This is a convenient way to add explanations or notes to your code without affecting its execution.

Formatting and organizing code within the script and document by using multiple levels of headings (`#`, `##`, `###`, `####`), you can create a hierarchical structure that visually outlines different sections of your code. This can make it easier to navigate and understand, especially for larger or more complex projects.

*Examples:*
```{r comment}
# This is a comment, Calculate the sum of 2 and 2
2 + 2

# This is a comment, Calculate the difference of 27 subtracted by 10
27 - 10
```

---

### **Basic Arithmetic**
R is a powerful tool for performing various mathematical calculations. It supports a wide range of arithmetic operations, including:

#### **Arithmetic Operators**
- Multiplication: `*`
- Division: `/`
- Addition: `+`
- Subtraction: `-`
- Exponentiation: `^`
- Modulo: `%%`

*Examples:*
```{r arithmetic-operators}

# Multiplication
# This is a comment, Calculate the product of 9 multiplied by 8
9 * 8

# Division
# This is a comment, Calculate the quotient of 6 divided by 2
6 / 2

# Addition
# This is a comment, Calculate the sum of 1 and 1
1 + 1

# Subtraction
# This is a comment, Calculate the difference of 27 subtracted by 10
27 - 10

# Exponential
# # This is a comment, Calculate 'yth power of x', so called 
# "The 5th power of 2 is 32"
2^5 # (2 * 2 * 2 * 2 * 2)

# Modulo
# # This is a comment, Calculate the remainder after division
# "28 modulo 6 is 4"
28%%6 # 28 divided by 6 is equal to 24, and remainder 4 is the modulo
```

---

### **Variables**
Variable is a container that stores a value, variable can hold a variety of data types such as numbers, text (strings), logical values (TRUE or FALSE) and others. <br>
`<-` assignment values using the assignment operator

*Examples:*
```{r variables}
# Assign the value of 28 to x
x <- 28

# Print out the value of the variable x
x

# Creating a variable named "age" and assigning the value of 38 to it
age <- 38

# Print out the value of the variable age using print built-in function
print(age)

### examples ----
# Assign a value to the variables my_salary and your_salary
my_salary <- 5000
your_salary <- 4500

# Add these two variables together
our_salary <- my_salary + your_salary

# Create the variable our_salary and print out
print(our_salary)
```

---

### **Data Types**
R, a versatile programming language for statistical computing and data analysis, recognizes several fundamental data types. Understanding these data types is crucial for effective data manipulation and analysis. Here's a breakdown of the primary data types in R:

*Examples:*
```{r data-types}
## Numeric ----
# Real numbers: Numbers with decimal points.
x <- 3.14159
print(x)

## Integers ----
# Whole Numbers.
y <- 42
print(y)

## Characters ----
# Strings: Sequences of characters enclosed in quotes.
text <- "Hello, world!"
print(text)

## Logical ----
# Boolean values: TRUE or FALSE. Represents 1 is TRUE and 0 is FALSE
is_true <- TRUE
is_false <- FALSE

print(is_true)
print(is_false)
```

---

### **class() function**
`class()` is a built-in function in R that is used to determine the data type of an object. It returns a character string specifying the class of the object.

*Examples:*
```{r class-function}
class(x)
class(y)
class(text)
class(is_true)
class(is_false)

```

---

### **str() function**
The `str()` function in R is a valuable tool for quickly inspecting the structure of an R object. It provides a concise summary of the object's type, dimensions, and a glimpse of its contents.

*Examples:*
```{r str-function}
str(x)
str(y)
str(text)
str(is_true)
str(is_false)

```