---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: false
    theme: cerulean
    code_folding: show
    
---

### What is matrix?
In R, the matrix() function is a versatile tool for creating matrices.<br>
A matrix is a rectangular array of numbers, symbols, or expressions, arranged in rows and columns.

```{r}
#matrix(data, nrow, ncol, byrow = FALSE, dimnames = NULL)
```
Arguements:<br>
• **data**: A vector containing the elements of the matrix.<br>
• **nrow**: The desired number of rows.<br>
• **ncol**: The desired number of columns.<br>
• **byrow**: A logical value indicating whether the matrix should be filled by row (TRUE) or by column (FALSE, default).<br>
• **dimnames**: A list of two character vectors, providing row and column names.<br>

---

### 3x3 Matrix
Create a 3x3 matrix with numbers 1 to 9, filled by row:

```{r matrix}
my_matrix <- matrix(1:9, nrow = 3, byrow = TRUE)
print(my_matrix)
```

---

### Analyze matrices
Box office movies of Star Wars named new_hope, empire_strikes and return_jedi
```{r analyze_matrix}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
print(box_office)

# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE)
print(star_wars_matrix)
```

---

### Naming a matrix
The `colnames()` function in R is a versitile tool used to: <br>

1. *Retrieve Column Names*: <br>
•  Returns a character vector containing the names of the columns in a data frame or matrix.

2. *Set Column Names*:<br>
•  Assigns new names to the columns of a data frame or matrix.<br>
The new names must be provided as a character vector of the same length as the number of columns.

The `rownames()` function in R is used to: <br>

1. *Retrieve Row Names*:<br>
•  Returns a character vector containing the names of the rows in a data frame or matrix.<br>

2. *Set Row Names*:<br>
•  Assigns new names to the rows of a data frame or matrix.<br>
The new names must be provided as a character vector of the same length as the number of rows.

```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)
print(star_wars_matrix)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
print(star_wars_matrix)
```

--- 

### Calculate: `rowSums()`
The `rowSums()` function in R is used to calculate the sum of values across each row of a matrix or data frame.
```{r rowSums}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", 
                 "The Empire Strikes Back", 
                 "Return of the Jedi")

# Create a matrix named as star_wars_matrix     
star_wars_matrix <- matrix(box_office, 
                      nrow = 3, byrow = TRUE,
                      dimnames = list(titles, region))

# Display the created matrix
print(star_wars_matrix)

# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Print the rowSums() result
worldwide_vector
```

---

### Adding Column: `cbind()`
The `cbind()` function in R is a versatile tool used to combine multiple objects (vectors, matrices, or data frames) by columns.<br>
This means it stacks the objects horizontally, creating a new object with the combined columns.
```{r cbind}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", 
            "The Empire Strikes Back", 
            "Return of the Jedi")
               
star_wars_matrix <- matrix(box_office, 
                      nrow = 3, byrow = TRUE,
                      dimnames = list(titles, region))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)
# Print worldwide_vector rowSums
print(worldwide_vector)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
# Print result of cbind() of start_wars_matrix and worldwide_vector
print(all_wars_matrix)
```

---

### Calculate: `colSums()`
`colSums()` in R: Summing Up Columns <br>

The `colSums()` function in R is a handy tool to calculate the sum of values within each column of a matrix or data frame. 

```{r all_wars_matrix}
# all_wars_matrix is available in your workspace
print(all_wars_matrix)

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)
  
# Print out total_revenue_vector
print(total_revenue_vector)
```

---


### Adding Row `rbind()`

The `rbind()` function in R is used to combine multiple objects (vectors, matrices, or data frames) by rows. This means it stacks the objects vertically, creating a new object with the combined rows.
```{r rbind}
### Adding Row: `rbind()` function
star_wars_vector <- c(461.0, 314.4, 290.5, 247.9, 309.3, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strike Back", "Return of the Jedi")

star_wars_matrix <- matrix(star_wars_vector,
                           nrow = 3, byrow = TRUE,
                           dimnames = list(titles, region))
print(star_wars_matrix)

star_wars_vector2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
titles2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")

star_wars_matrix2 <- matrix(star_wars_vector2,
                            nrow = 3, byrow = TRUE,
                            dimnames = list(titles2, region))
print(star_wars_matrix2)

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
# Print all_wars_matrix
print(all_wars_matrix)
```

---

### Selection of matrix elements
In R, the `matrix[]` syntax is a powerful tool for accessing and manipulating specific elements, rows, or columns within a matrix. <br>

`matrix[row_index, column_index]`

---

#### Subsetting a Row
`my_matrix[row_index, ]`

---

#### Subsetting a Column
`my_matrix[, column_index]`



```{r selection-matrix-elements}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
  
# Average non-US revenue
mean(non_us_all)
  
# Select the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2, 2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```

---

### Algorithm on matrix
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5
  
# Print the estimate to the console
visitors
```

---

### Arithmetic with matrices
```{r}

ticket_price <- c(5.0, 5.0, 6.0, 6.0, 7.0, 7.0, 4.0, 4.0, 4.5, 4.5, 4.9, 4.9)
star_wars_title <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi",
                     "The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
region <- c("US", "non-US")

ticket_prices_matrix <- 

ticket_prices_matrix <- matrix(ticket_price, nrow = 6, byrow = TRUE,
                               dimnames = list(star_wars_title, region))

                               
# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[,1]

# Average number of US visitors
mean(us_visitors)

```