---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: false
    theme: cerulean
    code_folding: show
    
---

---

### What is matrix?
In R, the matrix() function is a versatile tool for creating matrices.<br>
A matrix is a rectangular array of numbers, symbols, or expressions, arranged in rows and columns.

---
```{r}
#matrix(data, nrow, ncol, byrow = FALSE, dimnames = NULL)
```
Arguements:<br>
• **data**: A vector containing the elements of the matrix.<br>
• **nrow**: The desired number of rows.<br>
• **ncol**: The desired number of columns.<br>
• **byrow**: A logical value indicating whether the matrix should be filled by row (TRUE) or by column (FALSE, default).<br>
• **dimnames**: A list of two character vectors, providing row and column names.<br>

---

### 3x3 Matrix
Create a 3x3 matrix with numbers 1 to 9, filled by row:

```{r matrix}
my_matrix <- matrix(1:9, nrow = 3, byrow = TRUE)
print(my_matrix)
```

---

### Analyze matrices
Box office movies of Star Wars named new_hope, empire_strikes and return_jedi
```{r analyze_matrix}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
```
Combine the 3 vector into one vector named as box_office

```{r}
# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
print(box_office)
```

Create a matrix
```{r}
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE)
print(star_wars_matrix)
```