---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: false
    theme: cerulean
    code_folding: show
    
---

---

### What is matrix?
In R, the matrix() function is a versatile tool for creating matrices.<br>
A matrix is a rectangular array of numbers, symbols, or expressions, arranged in rows and columns.

---
```{r}
#matrix(data, nrow, ncol, byrow = FALSE, dimnames = NULL)
```
Arguements:<br>
• **data**: A vector containing the elements of the matrix.<br>
• **nrow**: The desired number of rows.<br>
• **ncol**: The desired number of columns.<br>
• **byrow**: A logical value indicating whether the matrix should be filled by row (TRUE) or by column (FALSE, default).<br>
• **dimnames**: A list of two character vectors, providing row and column names.<br>

---

### 3x3 Matrix
Create a 3x3 matrix with numbers 1 to 9, filled by row:

```{r matrix}
my_matrix <- matrix(1:9, nrow = 3, byrow = TRUE)
print(my_matrix)
```

---

### Analyze matrices
Box office movies of Star Wars named new_hope, empire_strikes and return_jedi
```{r analyze_matrix}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
```
Combine the 3 vector into one vector named as box_office

```{r}
# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)
print(box_office)
```

Create a matrix
```{r}
# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE)
print(star_wars_matrix)
```

---

### Naming a matrix
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)
print(star_wars_matrix)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
```
The `colnames()` function in R is a versitile tool used to: <br>

1. *Retrieve Column Names*: <br>
•  Returns a character vector containing the names of the columns in a data frame or matrix.

2. *Set Column Names*:<br>
•  Assigns new names to the columns of a data frame or matrix.<br>
The new names must be provided as a character vector of the same length as the number of columns.
```{r}
# Name the columns with region
colnames(star_wars_matrix) <- region
```
The `rownames()` function in R is used to: <br>

1. *Retrieve Row Names*:<br>
•  Returns a character vector containing the names of the rows in a data frame or matrix.<br>

2. *Set Row Names*:<br>
•  Assigns new names to the rows of a data frame or matrix.<br>
The new names must be provided as a character vector of the same length as the number of rows.

```{r}
# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
star_wars_matrix
```

---

### Calculate: `rowSums()` function
The `rowSums()` function in R is used to calculate the sum of values across each row of a matrix or data frame.
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", 
                 "The Empire Strikes Back", 
                 "Return of the Jedi")

# Create a matrix named as star_wars_matrix     
star_wars_matrix <- matrix(box_office, 
                      nrow = 3, byrow = TRUE,
                      dimnames = list(titles, region))

# Display the created matrix
print(star_wars_matrix)

# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Print the rowSums() result
worldwide_vector
```