---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

### Vector
Vectors are one of the most fundamental data structures in R.
They are essentially ordered collections of elements, where all elements must be of the same data type.<br>
In R, the `c()` function is a fundamental tool used to combine values into a vector. A vector is a one-dimensional array that can hold elements of the same data type.

```{r}
# Declared variable of different types using vector
my_number <- c(1, 10, 15)
print(my_number)

my_character <- c("z", "x", "y")
print(my_character, str(my_character))

my_logical <- c(FALSE, TRUE, FALSE)
print(my_logical)
```

### Names
The `names()` function is used to get or set the names of elements within an object, such as a vector, list, or data frame.

#### Setting Names
To assign names to an object, you use the assignment operator `<-` with the `names()` function:

```{r}
# Unnamed vector as daily_sales
daily_sales <- c(100, 250, 75, 15, 20, 90, 120)

# Assign names Monday to Sunday to unnamed vector daily_sales
names(daily_sales) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Print result
print(daily_sales)
```

### Getting Names
To retrieve the names of an object, you simply use the `names()` function:

```{r}
names(daily_sales)
```

### Setting names using variables
```{r}
# Number of fingers in sequence
finger_number <- c(1, 2, 3, 4, 5)
# Names of fingers from Thumb to Pinky
fingers_names <- c("Thumb", "Index", "Middle", "Ring", "Pinky")

# Assign finger_numbers to finger_names
names(finger_number) <- fingers_names

print(finger_number)
```

### Calculate vectors
```{r}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)

# Calculate the sum of sales_burger and sales_lemonade
total_sales <- sales_burger + sales_lemonade

# Create a vector of days
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Assign names
names(total_sales) <- days_vector

# Print the total_sales with days_vector
print(total_sales)

```