---
title: "Introduction to R"
author: "BotnieData"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 4
    number_sections: false
    theme: cerulean
    code_folding: show
    
---

---

### What is vector?
Vectors are one of the most fundamental data structures in R.
They are essentially ordered collections of elements, where all elements must be of the same data type.<br>

In R, the `c()` function is a fundamental tool used to combine values into a vector. A vector is a one-dimensional array that can hold elements of the same data type. Combining series of elements in a vector

---

### Vector using `c()` function
In R, the `c()` function is a fundamental tool used to combine values into a vector. A vector is a one-dimensional array that can hold elements of the same data type.

```{r combine}
# Declared variable of different types using vector
my_number <- c(1, 10, 15)
print(my_number, str(my_number))

my_character <- c("z", "x", "y")
print(my_character, str(my_character))

my_logical <- c(FALSE, TRUE, FALSE)
print(my_logical, str(my_logical))
```

---

### Names
The `names()` function is used to get or set the names of elements within an object, such as a vector, list, or data frame.

#### Setting Names
To assign names to an object, you use the assignment operator `<-` with the `names()` function:

```{r names}
# Unnamed vector as daily_sales_2
daily_sales <- c(100, 250, 75, 15, 20, 90, 120)

# Assign names Monday to Sunday to unnamed vector daily_sales
names(daily_sales) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Print result
print(daily_sales)
```
You can also assign names directly when creating a vector using `=` (equal) sign
```{r setting_names}

# Unnamed vector as daily_sales
daily_sales_2 <- c("Monday" = 100, "Tuesday" = 250, "Wednesday" = 75, "Thursday" = 15, "Friday" = 20, "Saturday" = 90, "Sunday" = 120)

# Print result
print(daily_sales_2)

```
Using `=` equal sign assigning the variable


#### Getting Names
To retrieve the names of an object, you simply use the `names()` function:

```{r getting_names}
names(daily_sales)
```

---

### Setting names using variables
```{r setting_names_2}
# Number of fingers in sequence
finger_number <- c(1, 2, 3, 4, 5)
# Names of fingers from Thumb to Pinky
fingers_names <- c("Thumb", "Index", "Middle", "Ring", "Pinky")

# Assign finger_numbers to finger_names
names(finger_number) <- fingers_names

print(finger_number)
```

---

### Vectors with Math Operator
```{r vector with math operator}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)

# Calculate the sum of sales_burger and sales_lemonade
total_sales <- sales_burger + sales_lemonade
# Create a vector of days
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Assign names
names(total_sales) <- days_vector

# Print the total_sales with days_vector
print(total_sales)

```

---

### Vector using `sum()` function
```{r sum function}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector
print(sales_burger)

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector
print(sales_lemonade)

# Create a vector of days
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Total sales per day
sales_per_day <- sales_burger + sales_lemonade
print(sales_per_day)

# Calculate the sum of sales_burger and sales_lemonade
sum_sales_burger <- sum(sales_burger)
print(sum_sales_burger)

sum_sales_lemonade <- sum(sales_lemonade)
print(sum_sales_lemonade)

# Calculate the total_sales_in_a_week
total_sales <- sum(sum_sales_burger, sum_sales_lemonade)

# Print the total_sales with days_vector
print(total_sales)
```

---

### Comparing vectors
R provides several methods for comparing vectors, both element-wise and overall.

Element-wise Comparison
You can use comparison operators like `==`, `!=`, `<`, `>`, `<=`, and `>=` to compare elements in two vectors of the same length. The result is a logical vector of the same length, indicating the truth value of each comparison.

```{r comparing vectors}
# Create two numeric vectors
vector1 <- c(10, 20, 30, 40)
vector2 <- c(15, 20, 25, 40)

# Compare elements using ==
result <- vector1 == vector2
print(result)
```

---

### Selection

#### Selection on single element
Selects the second elements from each vector, which corresponds to Tuesday's sales for burger and lemonade, respectively.
```{r selection in single value}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector
print(sales_burger)

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector
print(sales_lemonade)

# Create a vector of days
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Define a new variable based on a selection
# Selection of single element under Tuesday
burger_tuesday <- sales_burger[2]
lemonade_tuesday <- sales_lemonade[2]

# Print result
print(burger_tuesday)
print(lemonade_tuesday)

```

---

#### Selection containing elements
Selects the first, third and fifth elements from both vectors, which correspond to Monday, Wednesday and Friday's sales for each vector respectively.
```{r selection containing elements}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector

# Assign variable of days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Define a new variable based on a selection elements (first, mid and last day of the weekdays)
fml_burger <- sales_burger[c(1, 3, 5)]
fml_lemonade <- sales_lemonade[c(1, 3, 5)]

# Print out fml_burger (first, mid and last day of the weekday)
print(fml_burger)

# Print out fml_lemonade (first, mid and last day of the weekday)
print(fml_lemonade)
```

---

#### Selection range elements
Select the first and fourth elements of vector which correspond to Monday to Thursday's Burget and Lemonade Sales
```{r selection range elements}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector

# Assign variable of days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Define a new variable based on a selection from 1st to 4th element
mth_burger <- sales_burger[1:4]
mth_lemonade <- sales_lemonade[1:4]

# Select the 1st and 4th element on mth_burger
print(mth_burger)

# Select the 1st and 4th element on mth_lemonade
print(mth_lemonade)
```

---

#### Selection range with calculation
Calculating Average Weekday Burger and Lemonade Sales
```{r selection with calculation}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector

# Assign variable of days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

# Define a new variable based on a selection from Weekdays
weekdays_avg_sales_burger <- sales_burger[1:5]
weekdays_avg_sales_lemonade <- sales_lemonade[1:5]

# Calculate the average sales of burgers from Weekdays
avg_sales_burger <- mean(weekdays_avg_sales_burger)
print(avg_sales_burger)

# Calculate the average sales of lemonade from Weekdays
avg_sales_lemonade <- mean(weekdays_avg_sales_lemonade)
print(avg_sales_lemonade)
```

---

#### Selection comparisson
```{r}
# Create a vector of sales_burger
sales_burger <- c(100, 250, 75, 15, 20, 120, 150)
names(sales_burger) <- days_vector

# Create a vector of sales_lemonade
sales_lemonade <- c(20, 30, 10, 25, 20, 50, 70)
names(sales_lemonade) <- days_vector

# Assign variable of days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

---

#### Burger ----
```{r}
# Which days did you make a sales more than 50?
sales_burger_above50 <- sales_burger > 50
print(sales_burger_above50)

# Which days did you make a sales more than 50?
sales_burger_below50 <- sales_burger < 50
print(sales_burger_below50)

# Prints only the days with sales above 50
print(days_vector[sales_burger_above50])

# Prints only the days with sales below 50
print(days_vector[sales_burger_below50])

```
#### Lemonade ----
```{r}
# Which days did you make a sales more than 50?
sales_lemonade_above50 <- sales_lemonade > 50
print(sales_lemonade_above50)

# Which days did you make a sales more than 50?
sales_lemonade_below50 <- sales_lemonade < 50
print(sales_lemonade_below50)

# Prints only the days with sales above 50
print(days_vector[sales_lemonade_above50])

# Prints only the days with sales below 50
print(days_vector[sales_lemonade_below50])

```
